% !TeX root = ../paper.tex
% !TeX encoding = UTF-8
% !TeX spellcheck = en_US

\section{Related Work}\label{sec:related_work}
  The concept of order dependencies was first introduced in the context of database systems by \citeauthor{ginsburg}~\cite{ginsburg} as \textit{point-wise ordering}.
  \citeauthor{ginsburg}'s definition specified that a set of columns orders another set of columns.

  \citeauthor{szlichta:fundamentals}~\cite{szlichta:fundamentals} later introduced another definition for order dependencies, which was used by the following research in this area \cite{consonni, langer, szlichta:discovery} and is also the basis of this work.
  It differentiates from the point-wise ordering by considering list of columns instead of sets.
  This leads to a lexicographical ordering of tuples, as by the \texttt{order by} operator in SQL.

%TODO go into more detail about how the algorithms work?
\citeauthor{langer} published the first algorithm to find all minimal order dependencies based on \citeauthor{szlichta:fundamentals}'s definition. 
It is called \texttt{order} and has a factorial runtime.\\
When publishing their own algorithm \texttt{fastod}, \citeauthor{szlichta:discovery} showed, that \citeauthor{langer}'s definition of minimality was incomplete. 
\texttt{fastod} has an exponential worst time complexity \citep{szlichta:discovery}.\\
When introducing the algorithm \texttt{ocddiscover}, \citeauthor{consonni} claimed, that a provided implementation of \texttt{fastod} did not find all order dependencies. 
\texttt{ocddiscover} has a factorial runtime but was shown to outperform \texttt{fastod} on some datasets. 
It is also capable of multithreating, which makes it easy to adapt to a distributed system.
This was the state of published research when we started on this project, which is why we based our own algorithm on \texttt{ocddiscover}. \\
Since then, Szlichta et al have published an Errata Note demonstrating an error in \citeauthor{consonni}'s proof of minimality and explaining, how \citeauthor{consonni} misinterpreted \texttt{fastod}'s results.
Since the goal of our seminar was the distribution of an already existing algorithm with a focus on reliability, however, and scalability, we decided to continue working on \texttt{ocddiscover}.
